---
title: "Rworksheet_Mabalina#4b"
author: "Kristian B. Mabalina"
date: "2024-10-29"
output:
  pdf_document: default
  html_document: default
---
# 1
```{r}
vector <-c(1,2,3,4,5)
m1 <-matrix(nrow =5, ncol = 5)

for(i in 1:5){
  for(j in 1:5){
     m1[i, j] <- vector[i] + vector[j]
  }
}
print(m1)
```
# 2
```{r}
num_rows <- 5
for (i in 1:num_rows) {
  cat(rep("*", i), "\n")
}
```
# 3
```{r}
#start <- as.integer(readline(prompt = "Enter the starting number for the Fibonacci sequence: "))
start <- 5
a <- 0
b <- 1
repeat {
  fib <- a + b
  a <- b
  b <- fib
  if (fib >= start) {
    cat(fib, " ")
  }
  if (fib >= 500) {
    break
  }
}
```
# 4
```{r}
shoes <- read.csv("4bnum3.csv")
shoes
```
#b
```{r}
femalesubset <- subset(shoes, Gender.1 == "F")
malesubset <- subset(shoes, Gender.1 == "M")
femalecount <- nrow(femalesubset)
malecount <- nrow(malesubset)
cat("Female:", femalecount, "\n")
cat("Male:", malecount, "\n")
```
#c
```{r}
gendercounts <- table(shoes$Gender.1)
barplot(gendercounts, 
        main = "Number of Males and Females in Household Data",
        xlab = "Gender",
        ylab = "Count",
        col = c("blue", "red"),
        legend = c("Female", "Male"),
        names.arg = c("Female", "Male"))
```
# 5
```{r}
expenses <- c(Food = 60, Electricity = 10, Savings = 5, Miscellaneous = 25)
percent_labels <- paste0(names(expenses), " - ", round(expenses / sum(expenses) * 100), "%")
pie(expenses, 
    labels = percent_labels, 
    col = c("green", "blue", "yellow", "red"), 
    main = "Dela Cruz Family Monthly Expenses")

```
# 6
```{r}
data(iris)
str(iris)
```
#b
```{r}
colMeans(iris[, 1:4])
```
#c
```{r}
species_counts <- table(iris$Species)
pie(species_counts, 
    main = "Species Distribution in Iris Dataset", 
    col = c("blue", "green", "red"),
    labels = paste(names(species_counts), "\n", species_counts))
legend("topright", 
       legend = names(species_counts), 
       fill = c("blue", "green", "red"))

```
#d
```{r}
setosa <- subset(iris, Species == "setosa")
versicolor <- subset(iris, Species == "versicolor")
virginica <- subset(iris, Species == "virginica")
tail(setosa)
tail(versicolor)
tail(virginica)
```
#e
```{r}
colors <- c("setosa" = "red", "versicolor" = "blue", "virginica" = "green")
symbols <- c("setosa" = 16, "versicolor" = 17, "virginica" = 18)
plot(iris$Sepal.Length, iris$Sepal.Width, 
     col = colors[iris$Species], 
     pch = symbols[iris$Species],
     main = "Iris Dataset", 
     sub = "Sepal Width and Length",
     xlab = "Sepal Length", 
     ylab = "Sepal Width")
legend("topright", legend = c("Setosa", "Versicolor", "Virginica"), 
       col = c("red", "blue", "green"), 
       pch = c(16, 17, 18))

```
#f
#--The data structure proved suitable for both initial exploratory analysis and in-depth statistical modeling.
#--Mean values offered a quick snapshot of the key characteristics of the iris flowers.
#--The pie chart effectively showcased the species distribution, highlighting setosa as the dominant species.
#--Subsetting the dataset enabled a closer examination of each species, aiding in focused, species-specific analysis.
#--The scatterplot revealed the correlation between sepal length and width, visually distinguishing species and allowing further investigation into their relationships.

# 7
```{r}
library(readxl)
alexa_data <- read_excel("alexa_file.xlsx")
unique(alexa_data$variation)
alexa_data$variation <- gsub("Black Dot", "BlackDot", alexa_data$variation)
alexa_data$variation <- gsub("Black Plus", "BlackPlus", alexa_data$variation)
alexa_data$variation <- gsub("Black Show", "BlackShow", alexa_data$variation)
alexa_data$variation <- gsub("Black Spot", "BlackSpot", alexa_data$variation)
alexa_data$variation <- gsub("White Dot", "WhiteDot", alexa_data$variation)
alexa_data$variation <- gsub("White Plus", "WhitePlus", alexa_data$variation)
alexa_data$variation <- gsub("White Show", "WhiteShow", alexa_data$variation)
alexa_data$variation <- gsub("White Spot", "WhiteSpot", alexa_data$variation)
unique(alexa_data$variation)
```
#b
```{r}
library(dplyr)
variation_counts <- alexa_data %>%
  count(variation)
print(variation_counts)
save(variation_counts, file = "variations.RData")
```
#c
```{r}
library(dplyr)
load("variations.RData")
variation_counts$variation <- gsub(" +", " ", variation_counts$variation)
variation_counts$variation <- trimws(variation_counts$variation)
bar_data <- variation_counts$n
bar_names <- variation_counts$variation
barplot(
  bar_data,
  main = "Counts of Variations",               
  col = rainbow(length(bar_data)),              
  names.arg = bar_names,                        
  xlab = "Variations",                          
  ylab = "Counts",                            
  las = 2,                                      
  border = "black"                               
)
text(x = seq_along(bar_data), y = bar_data, labels = bar_data, pos = 3, cex = 0.8, col = "black")

```
#d
```{r}
library(ggplot2)
library(dplyr)
load("variations.RData")
variation_counts$variation <- gsub(" +", " ", variation_counts$variation)
variation_counts$variation <- trimws(variation_counts$variation)
bw_variations <- variation_counts %>%
  filter(grepl("Black|White", variation))
bar_data <- as.matrix(bw_variations$n)
bar_names <- bw_variations$variation
barplot(
  bar_data,
  beside = TRUE,                           
  main = "Counts of Black and White Variations", 
  col = c("black", "gray", "lightgray", "white"),  
  names.arg = bar_names,                   
  xlab = "Variations",                      
  ylab = "Counts",                          
  las = 2,                                
  border = "black"                          
)
text(x = seq_along(bar_data), y = bar_data, labels = bar_data, pos = 3, cex = 0.8, col = "black")

```


